---
title: "p8105_hw3_mt3866"
output: github_document
date: "2025-10-03"
---

##Question 1

Load the data:
```{r}
library(tidyverse)
library(dplyr)
library(p8105.datasets)
data("instacart")
```

Describe the data:
This dataset has 1,384,617 observations and 15 variables. In this dataset, each individual has a single order demonstrating items they bought using Instacart. Some key variables are reordered, which tells us how many times an individual reordered a certain item, order_id, which identifies the order number related to the data, product_id which can translate to which product they ordered, and user_id which identifies which user's data we are observing. For example, in one row user 112108 ordered product 11109 under order_id 1 and had reordered it before.

Identify number of aisles and aisles with greatest number of ordered items:
```{r}
instacart_df = instacart |>
count(aisle)|>
arrange(desc(n))
```


Filter to aisles with over 10,000 items ordered and then create ggplot:
```{r}
instacart_df_filtered = instacart_df |>
filter (n > 10000)

ggplot(instacart_df_filtered, aes(x = n, y = aisle)) +
  geom_point ()

```

Create a table of 3 most popular items in three aisles:
```{r}
instacart_popular = instacart |>
  filter (aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  count(aisle, product_name)|>
  group_by(aisle)|>
  arrange(desc(n))|>
  slice_head(n=3)
```

Bullet points:
There are 134 aisles. The aisles most ordered from are fresh vegetables, fresh fruits, packaged vegetables fruits, and yogurt. The three most popular items purchased in the aisle "baking ingredients" was light brown sugar (499), pure baking soda (387), and cane sugar (336). The three most popular items purchased in the aisle dog food care were snack sticks chicken & rice recipe dog treats (30), organix chicken & brown rice recipe (28), and small dog biscuits (26). The three most popular items ordered in packaged vegetables fruits were organic baby spinach (9,784), organic raspberries (5,546), and organic blueberries (4,966).